REDO Simulating Dwarf Galaxy Evolution with Individual Star Feedback and Chemical Evolution I: Methods and Test Case

### 2017-11-21 main.tex
into 4.2

\begin{figure*}
\centering
\includegraphics[width=0.975\linewidth]{multiplot_4x4_x.png}
\caption{textbf{Draft figure - need to pretty up and finalize fields and times.} Edge-on views of our dwarf galaxy at four different times in its evolution, beginning with the time at first star formation and ending at XXX Myr. Shown are density weighted projections of number density (top row), temperature slices (second row), and integrated projections giving \ion{H}{1} column density (third row) and O column density (fourth row). Additionally, each individual main sequence star particle is shown in the number density projections as a single black dot.}
\label{fig:panel_x}
\end{figure*}

The gas disk is is thin outside the stellar radius of our galaxy, but is resolved at all radii. In Figure~\ref{fig:scale_height} we plot the scale height of our gas disk as a function of radius averaged over the last 20 Myr of simulation time (black), at the initial conditions (blue), and just before first star formation (orange). At the outer edges of the disk, beyond regions of ongoing star formation, the scale height is $\sim 15$ pc, comparable to the scale height in the inner 200 pc prior to first star formation. Though small, both of these scale heights are resolved at our maximum resolution, 1.8 pc; we note that without feedback, our gas disk would continue to cool and collapse to an uresolved, razor thin disk. For the majority of the galaxy, stellar feedback sets the scale height. Heating, from both stellar radiation and supernovae, lifts gas above the midplane and generating scale heights between 40 and 100 pc within the half light radius. Effective galactic winds have carved out most of the gas within the inner 40 pc, distributing it somewhat smoothly above/below the midplane, leading to the significantly larger scale heights. This is made clear in the density panels if Fig~\ref{fig:panel_x}, where regions of active star formation show a general lack of the dense, cool gas that creates the smaller scale heights at larger radii.
%We additionally compute the mass weighted height of the disk, defined as $H = (\sum{Mz^2}/\sum{M})^{1/2}$, where the summation is over each gas cell, as a function of radius (not shown here). $H$ is generally larger than the scale height by a factor of 2, but is comparable to the scale height at radii with ongoing star formation, indicating a particularly uniform-then-truncated density distribution. At large radii, away from active star formation, $H$ is roughly 50 pc, or $\sim$ 3 times the scale height.

@@

Interestingly, there is significant structure well above the gas disk in the form of dense clouds and filamentary features, as seen in the edge-on projections. These gas cold gas clumps appear to have typical sizes on order of $\sim$ 10 pc, but with lower density tails that can extend up to 100 pc. These consist of both cool, self-gravitating gas clumps that are being pushed out and ablated by the stellar feedback driven winds, and gas that is cooling and condensing in the disk-halo interface of our dwarf galaxy. These clumps have typical velocities on order of a few to tens of km s$^{-1}$, and appear to survive for 10 - 100 Myr before being fully ablated or re-accreted onto the main body of the gaseous disk. The evolution of these dense clumps in the context of the disk-halo interface in our low mass dwarf galaxy will be examined in a future work. This is exciting given recent \citep[e.g.][]{Zheng2017} and ongoing \citep{Zheng2017HST} observational observational work examining the disk-halo interface in nearby galaxies.

@@

\subsection{SFR and Mass Evolution}
\label{sec:sfr}

We present the star formation rate (SFR) and core collapse supernova rate (SNR) evolution of our dwarf galaxy as measured in 10 Myr bins in the left panel of Figure~\ref{fig:sfr_mass_evolution}. By the end of the simulation, there have been no Type 1a supernovae. Within the first 40 Myr of evolution the SFR rises quickly to between $2-4 \times 10^{-4}$ M$_{\odot}$ yr$^{-1}$, and remains at this level with fluctuations of about a factor of 2 or less.\footnote{The granularity in our star formation algorithm creates a lower limit in the SFR that depends on the period over which the SFR is measured ($\Delta t$). Since we produce stars in $\sim 100$ M$_{\odot}$ sets, the smallest value for our measured SFR is $\sim 100/ \Delta t$. For $\Delta t = 10$ Myr this is 10$^{-5}$ yr$^{-1}$. Removing the granularity requires a fundamental change in our star formation algorithm, likely at the cost of increased complexity and computational expense. Sink particles, which track pre-main sequence stellar mass accumulation, would be the most viable way to do this \citep[see for example ][]{Krumholz2004,Federrath2010,GongOstriker2013,BleulerTeyssier2014,Sormani2017}.} Averaging over the entire simulation time since the first star was formed, we obtain a $<\rm{SFR}> = 2.34 \times 10^{-4}$ M$_{\odot}$ yr$^{-1}$. The star formation rate of our model galaxy as compared to observed galaxies is discussed further in Section~\ref{sec:observation}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.475\linewidth]{sfr_snr}
\includegraphics[width=0.475\linewidth]{mass_evolution}
\caption{Left: The SFR and core collapse SNR in our dwarf galaxy in 10 Myr bins (left). Right: The evolution of the total gas mass (black), \ion{H}{1} mass (blue), H$_2$ mass (orange), and stellar mass (red) in the galaxy over time. Note that the vertical axis limits in the left panel have been scaled to the same numerical values, yet with different units for the SFR and SNR.}
\label{fig:sfr_mass_evolution}
\end{figure*}

@@

The mass of the ISM in our dwarf galaxy is dominated by the CNM over the entire simulation time, as shown in the left panel of Figure~\ref{fig:ISM_evolution}. Together, the CNM and WNM comprise roughly 90\% of the galaxy's mass throughout most of its evolution. While the CNM has a negligible volume fraction, the WNM, WIM, and HIM occupy roughly the same volume each, though the two warm phases dominate. The volume fractions are subject to significantly more fluctuations over time, due to the low SFR and SNR of our galaxy. Together, these figures better quantify the general properties observed in the panel plots in Figure~\ref{fig:panel_x} and Figure~\ref{fig:panel_z}. These results are in some contrast with those found for the more massive dwarf galaxy in \citet{Hu2016,Hu2017}. In short, they find the mass and volume fraction of the ISM nearly entirely dominated by warm gas (ionized and neutral), with cold gas comprising $\sim$ 10\% of mass, and negligible volume, and hot gas comprising $\sim$ 10\% of the volume, but negligible mass. Our lower mass, lower metallicity galaxy contains more cold gas (by mass fraction) by a large factor, and more hot gas (by volume fraction) by a factor of a few.

>>>>

\begin{figure*}
\centering
\includegraphics[width=0.975\linewidth]{multiplot_4x4_x.png}
\caption{\textbf{Draft figure - need to pretty up and finalize fields and times.} Edge-on views of our dwarf galaxy at four different times in its evolution, beginning with the time at first star formation and ending at XXX Myr. Shown are density weighted projections of number density (top row), temperature slices (second row), and integrated projections giving \ion{H}{1} column density (third row) and O column density (fourth row). Additionally, each individual main sequence star particle is shown in the number density projections as a single black dot.}
\label{fig:panel_x}
\end{figure*}

The gas disk is thin outside the stellar radius of our galaxy, but is resolved at all radii. In Figure~\ref{fig:scale_height} we plot the scale height of our gas disk as a function of radius at the initial conditions (blue), just before first star formation (orange), and averaged over the last 20 Myr of simulation time (black). At the outer edges of the disk, beyond regions of ongoing star formation, the scale height is $\sim 15$ pc, comparable to the scale height in the inner 200 pc prior to first star formation. Though small, both of these scale heights are resolved at our maximum resolution, 1.8 pc; we note that without feedback, our gas disk would continue to cool and collapse to an uresolved, razor thin disk. For the majority of the galaxy, stellar feedback sets the scale height. Heating, from both stellar radiation and supernovae, lifts gas above the midplane and generating scale heights between 40 and 100 pc within the half light radius. Effective galactic winds have carved out most of the gas within the inner 40 pc, distributing it somewhat smoothly above/below the midplane, leading to the significantly larger scale heights. This is made clear in the density panels if Fig~\ref{fig:panel_x}, where regions of active star formation show a general lack of the dense, cool gas that creates the smaller scale heights at larger radii.
%We additionally compute the mass weighted height of the disk, defined as $H = (\sum{Mz^2}/\sum{M})^{1/2}$, where the summation is over each gas cell, as a function of radius (not shown here). $H$ is generally larger than the scale height by a factor of 2, but is comparable to the scale height at radii with ongoing star formation, indicating a particularly uniform-then-truncated density distribution. At large radii, away from active star formation, $H$ is roughly 50 pc, or $\sim$ 3 times the scale height.

@@

Interestingly, there is significant structure well above the gas disk in the form of dense clouds and filamentary features, as seen in the edge-on projections. These gas cold gas clumps appear to have typical sizes on order of $\sim$ 10 pc, but with lower density tails that can extend up to 100 pc. These consist of both cool, self-gravitating gas clumps that are being pushed out and ablated by the stellar feedback driven winds, and gas that is cooling and condensing in the disk-halo interface of our dwarf galaxy. These clumps have typical velocities on order of a few to tens of kilometers per second, and appear to survive for 10--100 Myr before being fully ablated or re-accreted onto the main body of the gaseous disk. The evolution of these dense clumps in the context of the disk-halo interface in our low mass dwarf galaxy will be examined in a future work. This is exciting given recent \citep[e.g.][]{Zheng2017} and ongoing \citep{Zheng2017HST} observational observational work examining the disk-halo interface in nearby galaxies.

@@

\subsection{Star Formation Rate and Mass Evolution}
\label{sec:sfr}

We present the star formation rate (SFR) and core collapse supernova rate evolution of our dwarf galaxy as measured in 10 Myr bins in the left panel of Figure~\ref{fig:sfr_mass_evolution}. By the end of the simulation, there have been no Type 1a supernovae. Within the first 40 Myr of evolution the SFR rises quickly to between $2-4 \times 10^{-4}$ M$_{\odot}$ yr$^{-1}$, and remains at this level with fluctuations of about a factor of 2 or less.  Averaging over the entire simulation time since the first star was formed, we obtain a $<\rm{SFR}> = 2.34 \times 10^{-4}$ M$_{\odot}$ yr$^{-1}$. The star formation rate of our model galaxy as compared to observed galaxies is discussed further in Section~\ref{sec:observation}.

%mm \footnote{  [This is rather too long for a footnote, I think.]
The granularity in our star formation algorithm creates a lower limit to the SFR that depends on the period over which the SFR is measured $\Delta t$. Since we produce stars in $\sim 100$~M$_{\odot}$ sets, the smallest value for our measured SFR is $\sim 100/ \Delta t$. For $\Delta t = 10$ Myr this is 10$^{-5}$ yr$^{-1}$. Removing the granularity requires a fundamental change in our star formation algorithm, likely at the cost of increased complexity and computational expense. Sink particles, which track pre-main sequence stellar mass accumulation, would be the most viable way to do this \citep[see for example ][]{Krumholz2004,Federrath2010,GongOstriker2013,BleulerTeyssier2014,Sormani2017}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.475\linewidth]{sfr_snr}
\includegraphics[width=0.475\linewidth]{mass_evolution}
\caption{Left: The SFR and core collapse supernova rate in our dwarf galaxy in 10 Myr bins (left). Right: The evolution of the total gas mass (black), \ion{H}{1} mass (blue), H$_2$ mass (orange), and stellar mass (red) in the galaxy over time. Note that the vertical axis limits in the left panel have been scaled to the same numerical values, yet with different units for the SFR and supernova rate.}
\label{fig:sfr_mass_evolution}
\end{figure*}

@@

The mass of the ISM in our dwarf galaxy is dominated by the CNM over the entire simulation time, as shown in the left panel of Figure~\ref{fig:ISM_evolution}. Together, the CNM and WNM comprise roughly 90\% of the galaxy's mass throughout most of its evolution. While the CNM has a negligible volume fraction, the WNM, WIM, and HIM occupy roughly the same volume each, though the two warm phases dominate. The volume fractions are subject to significantly more fluctuations over time, due to the low SFR and supernova rate of our galaxy. Together, these figures better quantify the general properties observed in the panel plots in Figure~\ref{fig:panel_x} and Figure~\ref{fig:panel_z}. These results are in some contrast with those found for the more massive dwarf galaxy in \citet{Hu2016,Hu2017}. In short, they find the mass and volume fraction of the ISM nearly entirely dominated by warm gas (ionized and neutral), with cold gas comprising $\sim$ 10\% of mass, and negligible volume, and hot gas comprising $\sim$ 10\% of the volume, but negligible mass. Our lower mass, lower metallicity galaxy contains more cold gas (by mass fraction) by a large factor, and more hot gas (by volume fraction) by a factor of a few.
